---
title: "Happy Analysis"
author: "Angela Lee"
date: "1/24/2017"
output: 
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: no
fontsize: 12pt
---



## Introduction
The 2018 World Happiness Report is a survey that ranks 162 countries by their happiness level. It is published by the Sustainable Development Solutions Network(SDSN). The 2018 release of Gallup World Poll covers 1,562 observations of happiness scores in the period of 2005 to 2017. Not all the countries and territories appear in all the years. For example, Canada is not included in the 2017 Happiness Index. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stats)

suppressWarnings(library(knitr))
suppressWarnings(library(readr))
Happiness <- read_csv("./WHR2018.csv")

```

##  Subset the data for using only the key features mentioned the World Happiness Report. 
They are GDP, Social Support, Healthy Life Expectancy, Freedom, Generosity, Perceptions of Corruption for the year between 2015 and 2016. The new data frame is called H1018.
```{r, echo= FALSE}

H2018 <- Happiness[, 1:9]
H2018 <- H2018 %>% filter(year == c(2015, 2016, 2017))

```


## Check the new dimension for H2018 - 158 x 9 

```{r}
dim(H2018)
```

## Checking Missing Data 

Rhere are 26 missing records in the new data frame - H2018. Fill missing data with 0's.

```{r, echo= FALSE}

sum(is.na(H2018))

H2018[is.na(H2018)] <- 0

sum(is.na(H2018))
```


## Happiness Score Distribution

```{r, echo= FALSE}
hist(H2018$`Life Ladder`, xlab = "Happiness Index", main = "Happiness Score Distribution - 2015, 2016, and 2018", col = "thistle")
```

The standard deviation is 1.03
```{r, echo= FALSE}
sd(H2018$`Life Ladder`)
```


## Create a function for the linear data normalization

```{r, include= TRUE}

normalization <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}
```


## Create a function for denormalization 
```{r, include= TRUE}

Unnormalize <- function(x, y) {
  y * (max(x) - min(x)) + min(x)
}
```


## Normalize the new dataframe - H2018

```{r, include=TRUE}
happy_df <- normalization(H2018[, 3:9])
sum(is.na(happy_df))
```             

## Let's create a multiple linear regression model using the six key features. 


We look at the cofficients in the t-test. The MLR function for the six key features is the following:

Life Ladder (Happiness) = -0.04 + 0.03 GDP + 1.16 Social Support + 0.09 Life Expectancy + 1.57 Freedom - 0.00 Generosity - 0.67 Corruption

The coefficient of multiple determination, R squared lies between 0 and 1 and is a descriptive measure of the utility of the regression for making predictions.

The adjusted for R squared is 0.61  ($R^{2}$= 0.61), which indicates the regression equation is SOMEWHAT useful for making predictions.
```{r, echo=FALSE}
df <- lm(`Life Ladder` ~., data = happy_df)
summary(df)
```

## Model Intepretation:
Our best estimate for the common standard deviation of all life ladder (happiness index) for all countries at any particlar GDP, Social Support, Healthy Life Expectancy, Freedom, Generosity, and corruption is 0.8674. 

## Variance Inflation Factor: 

Let's check if any of the six key variables have multicollinearity. VIF measures and indicates how much variance of an estimated regression coefficient is increased because of collinearity. If any of these variables has high VIF, then it is a highly correlated predictor variable. By looking at the variance inflation factor below, none of these key variables has VIF higher than 5, thus they are not highly correlated predictor variables. 

```{r, echo= FALSE}
library(car)
warning = FALSE
vif(df)
```


## ANNOVA - MLR Hypothesis case 1. 

We test if all slope parameters are 0.

The following is the hypothesis for the mutilple linear regression model for six features.


$$   \begin{aligned} H_0: \beta_1 GDP + \beta_2 SocialSupport+ \beta_3 Health + \beta_4 Freedom + \beta_5 Generosity + \beta_6 PerceptionCorruption = 0 \end{aligned} $$
$$   \begin{aligned} H_a:  At\ Least\ One\ Of\ The\ Variables\ \beta_j \not= 0\end{aligned} $$
Full Model: 
$$   \begin{aligned} Life\ Ladder\ (Happiness\ Index) = \beta_0 + \beta_1 GDP + \beta_2 SocialSupport+ \beta_3 Health + \beta_4 Freedom + \beta_5 Generosity + \beta_6 Perception\ Corruption \end{aligned}  $$
Reduced Model:
$$   \begin{aligned} Life\ Ladder\ (Happiness\ Index) = \beta_0 \end{aligned}$$
## Hypothesis Intepretation:
By looking at the F statistics and the p value (${p < 2.2e-16 }$), we can interpret that at the 5% significance level, the data provide sufficient evidence to conclude that at least one of the life ladder (happiness index) regression coefficients is not 0. Therefore, taken together, GDP, social support, healthly life expectancy at birth, freedom, generosity, and perceptions of corruption are useful in predicting happiness for a country. 

## ANNOVA - MLR Hypothesis case 2. 

By looking at the p value for the t test above, it looks like GDP, Social Support, and Generosity are not statistically significant. In other words, is happiness significantly related to GDP, Soical Support, and Generosity after taking into account other factors such as Health, Freedom, and Perceptions of Corruption? Let's make another hypothesis for the second case.  

$$   \begin{aligned} H_0: \beta_1 GDP = \beta_2 SocialSupport = \beta_5 Generosity  = 0 \end{aligned} $$
$$   \begin{aligned} H_a: At\ Least\ One\ Of\ The\ Variables\  \beta_1 , \beta_2 ,\beta_5 \not= 0 \end{aligned} $$
Full Model: 
$$   \begin{aligned} Life\ Ladder\ (Happiness\ Index) = \beta_0 + \beta_1 GDP + \beta_2 SocialSupport+ \beta_3 Health + \beta_4 Freedom + \beta_5 Generosity + \beta_6 Perception\ Corruption \end{aligned}  $$
Reduced Model:
$$   \begin{aligned} Life\ Ladder\ (Happiness\ Index) = \beta_0 + \beta_3 Health + \beta_4 Freedom + \beta_6 Perception\ Corruption \end{aligned}  $$
## ANNOVA for the partial model interpretation: 

By looking at the p value for the f test (p > 0.5), we do not reject the null hypothesis. At the 5% significance level, the data doesn't provide sufficient edvidence to reject the null ${beta_1 = beta_2 = beta_5 = 0}$ . Hence, in conjunction with other variables, GDP, Social Support, and Generosity are not useful predictors of the happiness. 
```{r, include= FALSE}
partial_H <- lm(`Life Ladder` ~ `Healthy life expectancy at birth` + `Freedom to make life choices` + `Perceptions of corruption` , data = happy_df)

anova(partial_H, df)

```


